---
title: "RegressionProject"
output: pdf_document
---

#Executive Summary
Based on the model created in this analysis, take MPG as outcome, the predictors include weight, number of cylinders, transmission type, and the interaction between weight and transmission type. Manual transmission has better MPG than auto. The MPG difference could be approximately quantified as: With weight and number of cylinders being equal, a car with manual transmission will have an improved MPG in the value of 11.9 minus 2.4 times the weight of car divided by 1,000, subtract the number of cylinders mutiplied by 1.2, further subtract 4.2 times the weight divided by 1,000, finally plus 34.3.  

##Introduction
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:  

- "Is an automatic or manual transmission better for MPG”  
- "Quantify the MPG difference between automatic and manual transmissions" 


Report Requirements:

Written as a PDF printout of a compiled (using knitr) R markdown document.
Brief. Roughly the equivalent of 2 pages or less for the main text. Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures.
Include a first paragraph executive summary.

Did the student interpret the coefficients correctly?
Did the student do some exploratory data analyses?
Did the student fit multiple models and detail their strategy for model selection?
Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
Did the student do a residual plot and some diagnostics?
Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
Was the report brief (about 2 pages long) for the main body of the report and no longer than 5 with supporting appendix of figures?
Did the report include an executive summary?
Was the report done in Rmd (knitr)?


## The Data
The data was extracted from the 1974 Motor Trend US magazine and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models). A table with more detailed information about each variable can be found in the appendix.

## Exploratory Analysis
The result of the Shapiro-Wilk normality test suggests it is safe to assume the dependent variable is normal with a mean of `r round(mean(mtcars$mpg),2)` mpg.

The exploratory plots suggest that manual transmission vehicles may be more fuel efficient than automatic transmission vehicles because the top row is slightly higher than the bottom row, but they also illustrate that other factors are probably effecting fuel efficiency.

For example, there seems to be an indirect relationship between fuel efficiency and weight.  This builds confidence in the quality of our data because it is reasonable to assume that heavier cars are less fuel efficient. A multi-variate method such as multi-linear regression will accurately assess how a car's transmission type (automatic/manual) effects its fuel efficiency while controlling for these other relationships.

## Full Analysis  

#### Exploratory Data Analysis  
First, load the data and examine the data strucutre (Figure 1). Looks like among 32 observations, it is fairely an even split between two groups in question (19 auto, 13 manual). Now create a boxplot to examine which one has better MPG on average (Figure 2). The boxplot indicates that manual transmission has better MPG than auto.    

#### Model Selection  
*Forward-Selection* model selection strategy is used here. First, create a basic model with a single variable: transmission type (am) (Figure 3).  

This model has an Estimate **p-value lower than 0.05**, thus proving there is a linear correlation between type of transmission to MPG.  

Next, we add one more predictor into this model. The next most likely one is weight (wt) (Figure 4).  Model 2's **coefficients all have low p-values**; it also has a better **Adjusted R-squared value**, suggesting both predictors need to be included in the final model.  

For Model 3, test if there is an interaction between transmission type and weight (Figure 5). Assumpetion is true. We also have reached a better **Ajusted R-squared value 0.815** while maintaining **low p-vlaues for all coefficients**.   

Now another predictor will be added: cylinder number (cyl) (Figure 6). **Low p-values, better Adjusted R-squared value as 0.853**. All the above predictors need to be selected. There also is no interaction between wt and cyl (Figure 7); but there is an interaction between cyl and am (Figure 8), however the **Adjusted R-squared value** is not as high as **0.853**.    

Adding the rest of the varaibles one by one and examine the coefficents p-values as well as if Adjusted R-squared improves, we cannot find any other predictors that improves the current model (process omitted from this analysis). Our final model is Model 4 in Figure 6:  

```{r}
fit4 <- lm(mpg ~ am*wt + cyl, mtcars)
```

#### Residuals examination  
Figure 9 exmmines the residuals vs. fitted values, as well as standardized residuals theoretical quantiles. The conclusions is there does not seem to be significant outliers other than three instances (the ones noted in the graph).  

#### Confidence Intervals  

```{r}
coef <- summary(fit4)$coef
coef <- data.frame(coef)
interval <- function(r) {
    r[1] + c(-1,1) * qt(.975, 27) * r[2]
}
t(apply(coef, 1, interval))
```


#### Answers to the Questions  

- "Is an automatic or manual transmission better for MPG”  
Because the coefficient for **am** is **positive (11.9385)**, and am takes 1 as the value for manual, **manual transmission** is better for MPG.  

- "Quantify the MPG difference between automatic and manual transmissions"  
With weight and number of cylinders being equal, a car with manual transmission will have an improved MPG in the value of **11.9385 minus 2.3689 times the weight of car divided by 1,000, subtract the number of cylinders mutiplied by 1.1814, further subtract 4.1974 times the weight divided by 1,000, finally plus 34.2830.** (refer to figure 6's coefficients)   


## Appendix

### Figure 1  
```{r, cache=TRUE, fig.width=6, fig.height=4}
head(mtcars, n=3)
## how many cobservations
nrow(mtcars)
## the distribution between two kinds of transmissions
table(mtcars$am)
```

### Figure 2  
```{r, cache=TRUE, fig.width=3, fig.height=2}
library(ggplot2)
ggplot(mtcars, aes(as.factor(am), mpg)) + geom_boxplot() + theme_bw() + labs(x="0=Auto, 1=Manual")
```

### Figure 3  

```{r}
fit1 <- lm(mpg ~ am, mtcars)
summary(fit1)$coef
summary(fit1)$adj.r.squared
```

### Figure 4  


```{r}
fit2 <- lm(mpg ~ am + wt, mtcars)
summary(fit2)$coef
summary(fit2)$adj.r.squared
```

### Figure 5  

```{r}
fit3 <- lm(mpg ~ am*wt, mtcars)
summary(fit3)$coef
summary(fit3)$adj.r.squared
```

### Figure 6 - FINAL MODEL    

```{r}
fit4 <- lm(mpg ~ am*wt + cyl, mtcars)
summary(fit4)$coef
summary(fit4)$adj.r.squared
```

### Figure 7  

```{r}
fit5 <- lm(mpg ~ am*wt + cyl*wt, mtcars)
summary(fit5)$coef
summary(fit5)$adj.r.squared
```

### Figure 8  
```{r}
fit6 <- lm(mpg ~ am*wt + cyl*am, mtcars)
summary(fit6)$coef
summary(fit6)$adj.r.squared
```

### Figure 9

```{r, fig.height=4,fig.width=8}
par(mfrow = c(1,2), mar = c(4, 4, 2, 1))
plot(fit4, which=1)
plot(fit4, which=2)
```
